Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2019-01-11 paper/SI.tex
fix typo in name
Add changes SI from Roman

The conductivity increases $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity versus concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 1.51$\pm$0.27$\times$10$^{-2}$  & 106   & 1.34 $\pm$ 0.24\\
    NaCl        & 9.90$\pm$0.17$\times$10$^{-3}$ & 91    & 1.02 $\pm$ 0.02\\
    LiCl        & 6.03 $\pm$0.80$\times$10$^{-3}$        & 83    & 0.682 $\pm$ 0.090\\
    CaCl$_{2}$  & 1.25$\pm$0.1$\times$10$^{-2}$    & 149   & 0.789 $\pm$ 0.063\\
    MgSO$_{4}$  & 8.80$\pm$1.2$\times$10$^{-3}$   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}]$ & 2.40$\pm$0.28$\times$10$^{-2}$   & 395   & 0.575 $\pm$ 0.063\\
    K$_{2}$SO4$_{4}$    & 1.98$\pm$0.20$\times$10$^{-2}$  & 258   & 0.722 $\pm$ 0.073\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\linewidth]{img/SI-eta-all.pdf}
  \caption{Measured $\eta$ values as function of $V_{\mathrm{G}}$ for
    different salts used in this work. The concentration in HCR is 0.1
    mM in all cases.}
  \label{fig:eta-all}
\end{figure}

@@

Next we seek the value between $R_{\mathrm{PCTE}}$ and
$R_{\mathrm{G}}^{0}$ (ungated resistance of PG), $\delta$. It is
pointed out that the scaling laws for $R_{\mathrm{PCTE}}$ and
$R_{\mathrm{G}}^{0}$ differ. For a long channel like PCTE, the
resistance is derived from Fick's law as\cite{O_Hern_2014}:
\begin{equation}
\label{eq:R-PCTE}
 R_{\mathrm{PCTE}} = \frac{1}{D} \frac{L}{\pi r_{\mathrm{PCTE}}^{2}}
\end{equation}
whereas $D$ is the diffusivity of the specie, $L$ is the length of
channel ($\sim$24 $\mathrm{\mu}$m), and $r_{\mathrm{PCTE}}$ is the
average radius of the PCTE channels ($\sim$400 nm). On the other hand,
the transport through 2D pore with high radius/thickness ratio falls
into the regime of the Sampson formula \cite{Roscoe_1949}, where the
entrance resistance dominates over the bulk resistance, with the
formula as \cite{Roscoe_1949,Celebi_2014}:
\begin{equation}
  \label{eq:sampson}
  R_{\mathrm{g}}^{0} = \frac{1}{D} \frac{3 \pi}{r_{\mathrm{G}}}
\end{equation}
We therefore estimate the value of $\delta$ as:
\begin{equation}
  \label{eq:delta-resistance}
  \delta
  = \frac{R_{\mathrm{PCTE}}}{R_{\mathrm{G}}^{0}}
  = {\displaystyle
    \frac{{\displaystyle \frac{L}
        {\pi r_{\mathrm{PCTE}}^{2}}}}
    {{\displaystyle \frac{3 \pi}{r_{\mathrm{G}}} \frac{1}{\gamma}}}}
  \approx 2.20
\end{equation}
Alternatively, we may also estimate the range of $\delta$ from
experimental data in main text \Fig{} \ref{main-fig:2}b using the
relation:
\begin{equation}
  \label{eq:exp-delta}
  \delta = \left(\frac{\boldsymbol{J}_{\mathrm{PCTE}}}{\boldsymbol{J}_{\mathrm{PG}}^{0}} -1 \right)^{-1}
\end{equation}
by assuming the validity of the circuit analog model for all salt
systems. We find that the $\delta$ value estimated from the simple
geometric model agrees well with the experimental results. As shown in
\Fig{}, the theoretical estimated value of $\delta$ (organe line) lies
within the margin of the majority of experimental data, which is
exceptionally good given the simplicity of the model used.

Finally we show how the effective salt rejection ratio through PG, $\xi$,
is extracted from the experimental $\eta$ values. Using the circuit
analog model, the definition of $\eta$ from main text
Eq. \ref{main-eq:rejection} can be rewritted using transport
resistances as:
\begin{equation}
  \begin{aligned}[t]
      \label{eq:eta-R}
  \eta &= 1 - {\displaystyle
    \frac{(R_{\mathrm{PCTE}} + R_{\mathrm{G}})^{-1}}
    {(R_{\mathrm{PCTE}} + R_{\mathrm{G}}^{0})}} \\
    &= 1 - \frac{\delta + 1}{\delta + R_{\mathrm{G}}/R_{\mathrm{G}}^{0}} \\
    &= 1 - \frac{\delta + 1}{\delta + \chi}
  \end{aligned}
\end{equation}
where we define $\chi = R_{\mathrm{G}}/R_{\mathrm{G}}^{0}$ which
represents the change of the transport resistance through PG upon
gating. Consequentially, the salt rejection ratio $\xi$, is associated
with $\chi$ by $\xi = 1 - \chi^{-1}$. Reformatting Eq. \ref{eq:eta-R}
gives:
\begin{equation}
  \label{eq:xi-deriv}
  \begin{aligned}[t]
    \xi &= 1 - \chi^{-1} \\
    &= \frac{(\delta + 1) \eta}{\delta\eta + 1}
  \end{aligned}
\end{equation}
which is main text Eq. \ref{main-eq:xi-def}. When $0<\eta<1$ (salt
rejection regime) and $\delta>0$, Eq. \ref{eq:xi-deriv} always
guarantees $\xi>\eta$, since $\delta + 1$ is always larger than
$\delta\eta + 1$. In other words, due to the existance of
$R_{\mathrm{PCTE}}$, the effective salt rejection through PG upon
gating, is always larger than that expperimentally observed.

@@

\section{Further details of the numerical simulation }
\label{sec:numer}

>>>>

The conductivity increase $s$ (mS$\cdot$cm$^{-1}\cdot$d$^{-1}$) is
converted into ion permeation rates $\boldsymbol{J}_{\mathrm{PCTE}}$
(mol$\cdot$m$^{-2}\cdot$s$^{-1}$) by use of a calibration factor
$c_{\mathrm{f}}$, the volume of LCR ($V_{\mathrm{LCR}}$) , the area of
the membrane $A$ and the time $t$ as:
\begin{equation}
  \label{eq:JPCTE}
  \boldsymbol{J}_{\mathrm{PCTE}} = \frac{s V_{\mathrm{LCR}}}{c_{\mathrm{f}} A t}
\end{equation}
The calibration factors for all salts have been obtained by a linear
fit of conductivity \textit{versus} concentration for 4 solutions with
concentrations from 10$^{-4}$ M to 10$^{-2}$ M and extracting the
resulting slope. The rest of the factors are $V_{\mathrm{LCR}}$ = 7.33
mL, $A$ = 9$\times$10$^{-6}$ m$^{-2}$. Table \ref{tab:exp-2} shows the
averaged conductivity increase s, the extracted calibration factors
$c_{\mathrm{f}}$ and the resulting ion permeation rates using
Eq. \ref{eq:JPCTE}. The values from this table are shown in main text \Fig 2b
(blue bars).
\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \hline
    Salt & $s$ (mS$\cdot$cm$^{-1}$d$^{-1}$) & $c_{\mathrm{f}}$ (mS$\cdot$cm$^{-1}\cdot$mol$^{-1}\cdot$ml) &
                                                                                                            $\boldsymbol{J}_{\mathrm{PCTE}}$ (10$^{6}$ mol$\cdot$m$^{-2}\cdot$s$^{-1}$)\\
    \hline
    KCl & 1.51$\pm$0.27$\times$10$^{-2}$  & 106   & 1.34 $\pm$ 0.24\\
    NaCl        & 9.90$\pm$0.17$\times$10$^{-3}$ & 91    & 1.02 $\pm$ 0.02\\
    LiCl        & 6.03 $\pm$0.80$\times$10$^{-3}$        & 83    & 0.682 $\pm$ 0.090\\
    CaCl$_{2}$  & 1.25$\pm$0.1$\times$10$^{-2}$    & 149   & 0.789 $\pm$ 0.063\\
    MgSO$_{4}$  & 8.80$\pm$1.2$\times$10$^{-3}$   & 130   & 0.633 $\pm$ 0.084\\
    K$_{3}$[Fe(CN)$_{6}]$ & 2.40$\pm$0.28$\times$10$^{-2}$   & 395   & 0.575 $\pm$ 0.063\\
    K$_{2}$SO4$_{4}$    & 1.98$\pm$0.20$\times$10$^{-2}$  & 258   & 0.722 $\pm$ 0.073\\
    \hline
  \end{tabular}
  \caption{Extracted slope $s$ as an averaged result from 3 consecutive
    measurements, the calibration factors $c_{\mathrm{f}}$ and the resulting ion
    permeation $\boldsymbol{J}_{\mathrm{PCTE}}$ through b-PCTE membranes for all salts.}
  \label{tab:exp-2}
\end{table}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.7\linewidth]{img/SI-eta-all.pdf}
  \caption{Measured $\eta$ values as function of $V_{\mathrm{G}}$ for
    different salts used in this work. The concentration in HCR is 0.1
    mM in all cases.}
  \label{fig:eta-all}
\end{figure}

@@

Next we seek the value between $R_{\mathrm{PCTE}}$ and
$R_{\mathrm{G}}^{0}$ (ungated resistance of PG), $\delta$. It is
pointed out that the scaling laws for $R_{\mathrm{PCTE}}$ and
$R_{\mathrm{G}}^{0}$ differ. For a long channel like PCTE, the
resistance is derived from Fick's law as\cite{O_Hern_2014}:
\begin{equation}
\label{eq:R-PCTE}
 R_{\mathrm{PCTE}} = \frac{1}{D} \frac{L}{\pi r_{\mathrm{PCTE}}^{2}}
\end{equation}
whereas $D$ is the diffusivity of the species, $L$ is the length of
channel ($\sim$24 $\mathrm{\mu}$m), and $r_{\mathrm{PCTE}}$ is the
average radius of the PCTE channels ($\sim$200 nm). On the other hand,
the transport through 2D pore with high radius/thickness ratio falls
into the regime of the Sampson formula \cite{Roscoe_1949}, where the
entrance resistance dominates over the bulk resistance, with the
formula as \cite{Roscoe_1949,Celebi_2014}:
\begin{equation}
  \label{eq:sampson}
  R_{\mathrm{g}}^{0} = \frac{1}{D} \frac{3 \pi}{r_{\mathrm{G}}}
\end{equation}
We therefore estimate the value of $\delta$ as:
\begin{equation}
  \label{eq:delta-resistance}
  \delta
  = \frac{R_{\mathrm{PCTE}}}{R_{\mathrm{G}}^{0}}
  = {\displaystyle
    \frac{{\displaystyle \frac{L}
        {\pi r_{\mathrm{PCTE}}^{2}}}}
    {{\displaystyle \frac{3 \pi}{r_{\mathrm{G}}} \frac{1}{\gamma}}}}
  \approx 2.20
\end{equation}
Alternatively, we may also estimate the range of $\delta$ from
experimental data in main text \Fig{} \ref{main-fig:2}b using the
relation:
\begin{equation}
  \label{eq:exp-delta}
  \delta = \left(\frac{\boldsymbol{J}_{\mathrm{PCTE}}}{\boldsymbol{J}_{\mathrm{PG}}^{0}} -1 \right)^{-1}
\end{equation}
by assuming the validity of the circuit analog model for all salt
systems. We find that the $\delta$ value estimated from the simple
geometric model agrees well with the experimental results. As shown in
\Fig{} \ref{}, the theoretical estimated value of $\delta$ (orange
line) lies within the margin of the majority of experimental data,
which is exceptionally good given the simplicity of the model used.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.75\linewidth]{img/SI-delta-estimate.pdf}
  \caption{Comparison between the $\delta$ value estimated from simple
    geometric model (Eq. \ref{eq:delta-resistance}) and from
    experimental value (Eq. \ref{eq:exp-delta}), showing good
    agreement between the two methods.  }
  \label{fig:delta-compare}
\end{figure}

Finally we show how the effective salt rejection ratio through PG, $\xi$,
is extracted from the experimental $\eta$ values. Using the circuit
analog model, the definition of $\eta$ from main text
Eq. \ref{main-eq:rejection} can be rewritted using transport
resistances as:
\begin{equation}
  \begin{aligned}[t]
      \label{eq:eta-R}
  \eta &= 1 - {\displaystyle
    \frac{(R_{\mathrm{PCTE}} + R_{\mathrm{G}})^{-1}}
    {(R_{\mathrm{PCTE}} + R_{\mathrm{G}}^{0})}} \\
    &= 1 - \frac{\delta + 1}{\delta + R_{\mathrm{G}}/R_{\mathrm{G}}^{0}} \\
    &= 1 - \frac{\delta + 1}{\delta + \chi}
  \end{aligned}
\end{equation}
where we define $\chi = R_{\mathrm{G}}/R_{\mathrm{G}}^{0}$ which
represents the change of the transport resistance through PG upon
gating. Consequentially, the salt rejection ratio $\xi$, is associated
with $\chi$ by $\xi = 1 - \chi^{-1}$. Reformatting Eq. \ref{eq:eta-R}
gives:
\begin{equation}
  \label{eq:xi-deriv}
  \begin{aligned}[t]
    \xi &= 1 - \chi^{-1} \\
    &= \frac{(\delta + 1) \eta}{\delta\eta + 1}
  \end{aligned}
\end{equation}
which is main text Eq. \ref{main-eq:xi-def}. When $0<\eta<1$ (salt
rejection regime) and $\delta>0$, Eq. \ref{eq:xi-deriv} always
guarantees $\xi>\eta$, since $\delta + 1$ is always larger than
$\delta\eta + 1$. In other words, due to the existance of
$R_{\mathrm{PCTE}}$, the effective salt rejection through PG upon
gating, is always larger than that experimentally observed.

@@

\pagebreak
\section{Further details of the numerical simulation }
\label{sec:numer}
